### 2019-10-16 thesis/txt/0-introduction.tex
writing about prose technique and some literature notes

\chapter{Introduction}
\plan{motivate why information extraction from build logs}
Many software projects now use continuous integration (CI)~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect bugs more easily~\cite{vasilescu2015quality,duvall2007continuous} and improve developer productivity~\cite{miller2008hundred,hilton2016usage} and communication~\cite{downs2012ambient}. These CI builds often produce very long and verbose build logs \mention{add log characteristics, cite what Moritz cited in his proposal?}, stating the progress and results of the various steps within the build.

These build logs are a highly valuable data source. First of all, for the developers that read them to analyze why their build failed or .. \mention{find more reasons}. Second, for researchers that can harvest the information contained in the logs -- and their metadata -- to study the software engineering process of a project. However they can only utilize the information within the build logs if they can adequately extract the information relevant to them.

There are many different extraction techniques use for this task. Beller et al.  used regular expressions when analysing the failure reasons of ruby and Java Maven buildlogs from TravisCI~\cite{beller2017oops}, while Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}. Recently Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}. \bp{anecdotal: keyword search?}. Apart from those there are various more extraction techniques like searching for keywords, ... \todo{more?} \bp{introduce problem: which one to choose}

With our work we want to support developers, researchers or project managers in deciding which technique is the best one for their use case.

\plan{what we do in this thesis:}

We aim to define a model to characterize the different extraction techniques and answer what influences the suitability of a technique. We support our assumption by evaluating three chosen techniques, namely regular expression synthesis by example using the Microsoft PROSE library, a common text similarity approach and simple keyword search. Our \emph{Failing Build Logs Data Set} encompasses about 800 log files from 80 repositories, labeled with the log part describing the reason a build failed, keywords to search for this extractions and a categorization of the extractions according to their structural position within the log \todo{is this understandable here? leave out?}.

We aim to answer the following research questions:

\begin{itemize}
  \item[\textbf{RQ1:}] What criteria influence the suitability of an information extraction technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to extract information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the extractions of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}

Our study of the three techniques on \emph{Failing Build Logs Data Set} shows that PROSE is suitable for extraction tasks requiring a high precision, ... . Extracting information by using text similarity of examples is suitable when ... . Simple keyword search is suitable ... .

\plan{contributions:}

Our work contributes:

\begin{itemize}
  \item A model to characterize information extraction techniques from \todo{CI?} build logs
  \item A model to characterize use case scenarios for information extraction from CI build logs
  \item A model of the extractable information in CI build logs
  \item A tool unifying several information extraction techniques namely:
        \begin{itemize}
          \item our implementation of regular expression program synthesis using the Microsoft PROSE library,
          \item a common information retrieval approach using text similarity, and
          \item a simple keyword search approach
        \end{itemize}
  \item The \emph{LogCollector}, a tool to gather logs from Travis CI
  \item A validated data set of about 800 logs from failed Travis CI builds labeled with:
        \begin{itemize}
          \item the substring of the log describing the reason the build failed,
          \item keywords developers would use to search for these descriptions of the build failure reason, and
          \item a categorization of the extractions according to their structural position within the build log
        \end{itemize}
  \item An evaluation of our model for the three implemented information extraction techniques \todo{describe \emph{what} we look for with this evaluation}
  \item A extendable sceme supporting decision on (our three) information extraction techniques relative to a given use case scenario
\end{itemize}

\plan{structure overview:}

>>>>

\chapter{Introduction}
%Many s These CI builds often produce very long and verbose build logs \mention{add log characteristics, cite what Moritz cited in his proposal?}, stating the progress and results of the various steps within the build.

%These build logs are a highly valuable data source. First of all, for the developers that read them to analyze why their build failed or .. \mention{find more reasons}. Second, for researchers that can harvest the information contained in the logs -- and their metadata -- to study the software engineering process of a project. However they can only utilize the information within the build logs if they can adequately extract the information relevant to them.

%There are many different extraction techniques use for this task. Beller et al.  used regular expressions when analysing the failure reasons of ruby and Java Maven buildlogs from TravisCI~\cite{beller2017oops}, while Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}. Recently Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}. \bp{anecdotal: keyword search?}. Apart from those there are various more extraction techniques like searching for keywords, ... \todo{more?} \bp{introduce problem: which one to choose}

%With our work we want to support developers, researchers or project managers in deciding which technique is the best one for their use case. 
Continuous Integration (CI) has become a best practice in software engineering and many software projects now use CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect bugs more easily~\cite{vasilescu2015quality,duvall2007continuous} and improve developer productivity~\cite{miller2008hundred,hilton2016usage} and communication~\cite{downs2012ambient}. A build on a CI server typically not only compiles and packages the software, but also executes test~\cite{beller2017oops} and various kinds of static analysis~\cite{zampetti2017open}.

CI builds often produce very long and verbose build logs~\cite{beller2017oops}, stating the progress and results of the various steps within the build. The structure of these logs changes greatly from project to project and highly depends on the tools and environment used.
Build logs are a highly valuable data source. First of all, for the developers that read them to analyze why their build failed or how different stages of the build process perform.
Second, for researchers that can harvest the information contained in the logs --- and their metadata --- to study the software engineering process of a project.
However these two groups can only utilize the information within the build logs if they can adequately extract the information relevant to them.

There are many different extraction techniques used for this task. Beller et al. used regular expressions when analyzing the failure reasons of Ruby and Java Maven build logs from TravisCI~\cite{beller2017oops}. Such regular expressions are developed by looking at a few exemplary build logs and maintaining them whenever new cases are introduced is a tedious and error-prone task~\cite{michael2019regexes}.
Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}.
Recently Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}.

Currently we only have anecdotal evidence on the performance of these techniques and when to use which. Developers and researches have little support while choosing which one to use for an extraction task. Our goal for this thesis is to investigate different information extraction techniques and describe under which circumstances we recommend each of the techniques.

We aim to define a model to characterize the different extraction techniques and answer what influences the suitability of a technique.
We support our assumption by evaluating three chosen techniques, namely regular expression synthesis by example using the Microsoft PROSE library (referred to as PBE), a common text similarity approach (referred to as TS) and simple keyword search (referred to as SKWS).
Our \emph{Failing Build Logs Data Set} encompasses about 800 log files from 80 repositories, labeled with the log part describing the reason a build failed, keywords to search for this extractions and a categorization of the extractions according to their structural position within the log.

We aim to answer the following research questions:
\begin{itemize}
  \item[\textbf{RQ1:}] What criteria influence the suitability of an information extraction technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to extract information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the extractions of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}

\plan{results and conclusions of my work}
Our study of the three techniques on \emph{Failing Build Logs Data Set} shows that PROSE is suitable for extraction tasks requiring a high precision, ... . Extracting information by using text similarity of examples is suitable when ... . Simple keyword search is suitable ... .

\paragraph{Our work contributes:}
\begin{itemize}
  \item A model of the extractable information in CI build logs
  \item A model to characterize information extraction techniques from CI build logs
  \item A model to characterize use case scenarios for information extraction from CI build logs
  \item A tool unifying several information extraction techniques namely:
        \begin{itemize}
          \item our implementation of regular expression program synthesis using the Microsoft PROSE library (PBE),
          \item a common information retrieval approach using text similarity (TS), and
          \item a simple keyword search approach (SKWS)
        \end{itemize}
  \item The \emph{LogCollector}, a tool to gather build logs from Travis CI
  \item A validated data set of about 800 logs from failed Travis CI builds labeled with:
        \begin{itemize}
          \item the substring of the log describing the reason the build failed,
          \item keywords developers would use to search for these descriptions of the build failure reason, and
          \item a categorization of the extractions according to their structural position within the build log
        \end{itemize}
  \item An evaluation of our model for the three implemented information extraction techniques \todo{describe \emph{what} we look for with this evaluation}
  \item A extendable scheme supporting decision on (our three) information extraction techniques relative to a given use case scenario
\end{itemize}

### 2019-10-16 thesis/txt/1-related-work.tex
writing on models chapter, putting pictures into figures, update label names
[wip] toc corrections
writing about prose technique and some literature notes

\chapter{Related Work}
\todo{introduction \& glue text}
This section presents various research works adjecent or foundational for our work. Starting with studies about Continuous Integration used in software projects, we move on to past works about build log analysis and their augmentation. We differentiate our work from the more common production runtime log analysis and cover different information extraction and retrieval techniques. Lastly this section mentions different Programming by Example resources surrounding Microsoft's work on the PROSE library.

\section{Continuous Integration Research}
\mention{look into proksch papers}
Various researchers have analyzed industrial and open source logs for failure reasons and their impact on development. Seo et al.~\cite{seo2014programmers} found that few error types such as dependency mismatches are the most prominent cause of build failures at Google. In addition most failures are resolved within two builds. Vassallo et al.~\cite{vassallo2017a-tale} compared open source projects in Java to industrial ones. They determined that testing failures outweigh compilation errors. Open source builds fail most often because of unit tests, whereas release preparations is the primary cause in industrial projects. Beller et al.~\cite{beller2017oops} showed that testing is central to continuous integration when evaluating Travis CI logs for Java and ruby builds. They observed very different kinds, failure rates and numbers of test between programming languages and explained that the low failure rates hint at code being tested before it is sent to the CI server.

@@

\section{Production Log Analysis}
\mention{logpai and drain}

>>>>

\chapter{Related Work}
\label{sec:rw}
\todo{introduction \& glue text}
This section presents various research works adjecent or foundational for our work. Starting with studies about Continuous Integration used in software projects, we move on to past works about build log analysis and their augmentation. We differentiate our work from the more common production runtime log analysis and cover different information extraction and retrieval techniques. Lastly this section mentions different Programming by Example resources surrounding Microsoft's work on the PROSE library.

\section{Continuous Integration}
\mention{look into proksch papers}
Various researchers have analyzed industrial and open source logs for failure reasons and their impact on development. Seo et al.~\cite{seo2014programmers} found that few error types such as dependency mismatches are the most prominent cause of build failures at Google. In addition most failures are resolved within two builds. Vassallo et al.~\cite{vassallo2017a-tale} compared open source projects in Java to industrial ones. They determined that testing failures outweigh compilation errors. Open source builds fail most often because of unit tests, whereas release preparations is the primary cause in industrial projects. Beller et al.~\cite{beller2017oops} showed that testing is central to continuous integration when evaluating Travis CI logs for Java and ruby builds. They observed very different kinds, failure rates and numbers of test between programming languages and explained that the low failure rates hint at code being tested before it is sent to the CI server.

@@

\section{Production Log Analysis}
\mention{logpai and drain, [7] K. Fisher and D. Walker. The pads project: an overview. In ICDT,
2011.
[8] K. Fisher, D. Walker, K. Q. Zhu, and P. White. From dirt to shovels:
fully automatic tool generation from ad hoc data. In POPL, 2008.[23] Q. Xi and D. Walker. A context-free markup language for semi- structured text. In PLDI, pages 221–232, 2010.
 The PADS project [7] has en- abled simplification of ad hoc data processing tasks for programmers
accounts addresses split chairs awk banks companies countries hadoop horses instruments ls-l mgx namephone nozzle numbertext papers pldi12 pldi13 pop13 quotes speechbench techfest ucla-faculty users
accounts addresses split chairs awk banks companies countries hadoop horses instruments ls-l mgx namephone nozzle numbertext papers pldi12 pldi13 pop13 quotes speechbench techfest ucla-faculty users
abt amazon apple barnes bestbuy bigtray bol buy cameraword cnet cooking-bw dealtime drugstore ebay mgzoutlet mediaworld nthbutsw powells googlepdct yahooshop shopping shopzilla target tigerdirect venere
abt amazon apple barnes bestbuy bigtray bol buy cameraword cnet cooking-bw dealtime drugstore ebay mgzoutlet mediaworld nthbutsw powells googlepdct yahooshop shopping shopzilla target tigerdirect venere
hg_ex12 hg_ex18 hg_ex2 hg_ex26 hg_ex29 hg_ex3 hg_ex39 _h8d62ck1* 03PFMJOU* 2003Fall 64040 anrep9899* bali ch15_e compliance* DataDiction* deliverable* e_Bubble_* flip_usd5 Funded - F ge-revenues HOSPITAL* pwpSurvey* SOA4-YEAR* young_table
hg_ex12 hg_ex18 hg_ex2 hg_ex26 hg_ex29 hg_ex3 hg_ex39 _h8d62ck1* 03PFMJOU* 2003Fall 64040 anrep9899* bali ch15_e compliance* DataDiction* deliverable* e_Bubble_* flip_usd5 Funded - F ge-revenues HOSPITAL* pwpSurvey* SOA4-YEAR* young_table
seconds
# examples
seconds
examples
seconds
examples
by contributing along several dimensions: development of domain specific languages for describing text structure or data format, learn- ing algorithms for automatically inferring such formats [8], and a markup language to allow users to add simple annotations to enable more effective learning of text structure [23] While PADS supports parsing of entire files, FlashExtract allows users to extract only parts of the file thereby avoiding unnecessary complications. PADS’s learner only supports a fixed line-by-line chunking strategy to split the records; in contrast, FlashExtract can learn chunking (aka, struc- ture boundaries) from examples, making it suitable for extracting data fields and records that have arbitrary length (and might cross multiple lines). Finally, PADS primarily targets ad hoc text files. Although one can view webpages and spreadsheet as text files, it is unclear if the PADS learning algorithm can be adapted to work effectively for webpages and spreadsheets.
}

### 2019-10-16 thesis/txt/2-models.tex
[wip] toc corrections
writing about prose technique and some literature notes

\section{Characteristics of a Build Log}
The idea of CI is, to integrate new software changes fast and often to catch errors early. After making a change, the developer commits and pushes it to a shared source code repository. Companies often use a specific CI server, e.g. Travis CI, linked to their source code repository. A CI build might then be triggered by a push on specific branches or after a pull request was created. When such a CI build is triggered, it typically runs through these stages:
\begin{itemize}
	\item Pulling in the new, changed version of the source code.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests.
\end{itemize}
However these are only \emph{typical} stages and there is a high variability in the CI build processes of different software projects \mention{citation needed, zampetti2017open?}. Some smaller projects might just use CI to make sure their code compiles as a minimal check before reviewing a pull request, other projects might have various stages of extensive automated testing.

@@

Semi-structured for build logs means for us that the log text is:
\begin{itemize}
	\item[implicitly stuructured], we do not have access to explicit structuring elements or an explicit structure description
	\item[irregular], the structure changes without notice \todo{we saw this in our data collection, random newlines}
\end{itemize}

@@

\subsection{Comparison to Runtime Logs}
\todo{or did we do that in related work enough?}

\section{Our Model for Extractable Information in Build Logs}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
  \caption{Information extractable from build logs}
  \label{fig:build-log-information-draft}
\end{figure}
Continuous integration build logs contain a huge amount of information about the various stages in the CI build they correspond to. In this section we present our model to more precisely describe our notion of a information extractable from a CI build log.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{img/mt-graphics-BuildLogInformation.pdf}
	\caption{Build Log Information}
	\label{fig:build-log-information}
\end{figure}
During our inital exploration \& data collection for the \emph{Failing Build Log Data Set}, wanted to get an impression about the information one would want to extract from a build log. Within the Travis CI build logs we found various different informations possibly interesting to be extracted
\begin{itemize}
	\item[Travis Build Step] A Travis CI build consists of several build steps defined within the Travis CI configuration language. Within the build log each of these steps are framed by \lstinline{travis_fold:start:<build step name>} and \lstinline{travis_fold:end:<build step name>}.
	      A build step contains
	      \begin{itemize}
		      \item[Build Step Name] The string Travis uses to identify the step within start \& end statements.
		      \item[Build Step Output] The output generated during the build step. The textual representation of the \texttt{Build Step Output} is the substring between the start \& end statements.
	      \end{itemize}
	\item[Travis Timing] Travis can measure the time of specified sections of the build process. It represents those by \lstinline{travis_time:start:<timing section id>} and \lstinline{travis_time:end:<timing section id>:start=<start time>,finish=<finish time>,duration=<duration>}
\end{itemize}

@@

\section{Our Model for Information Extraction Techniques}
\begin{figure}[h]
	\centering
	\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for an information extraction technique}
	\label{fig:model-ie-technique}
\end{figure}
\plan{model = making a picture with boxes to visualize what we are talking about}
A build log information extraction technique (BLIE technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
	\item{Target}{Extractable Information} each extraction technique \todo{instantiation!} targets a specific information
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
	\item \textbf{Configuration Scope} for one repository/project, a language/logkind, global
\end{itemize}

@@

\section{Our Model for Information Extraction Tasks}
\begin{figure}[h]
	\centering
	\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for information extraction task}
	\label{fig:modelt-ie-task}
\end{figure}
aims to describe an information extraction task of a developer or a researcher.
\begin{itemize}
	\item \todo{describe all classes}
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

\section{PROSE Program Synthesis}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Synthesizing regular expression programs by example using PROSE}
  \label{fig:prose-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?, describe regex programs}

\section{Text Similarity}
\begin{figure}[h]
  \centering
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using text similarity}
  \label{fig:text-similarity-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

>>>>

\section{Characteristics of a Build Log}
The idea of CI is, to integrate new software changes fast and often to catch errors early~\cite{humble2010continuous}. After making a change, the developer commits and pushes it to a shared source code repository. Companies often use a specific CI server, e.g. Travis CI, linked to their source code repository. A CI build might then be triggered by a push on specific branches or after a pull request was created. When such a CI build is triggered, it typically runs through these stages:
\begin{itemize}
	\item Pulling in the new, changed version of the source code.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests.
\end{itemize}
However these are only \emph{typical} stages and there is a high variability in the CI build processes of different software projects \mention{citation needed, zampetti2017open?}. Some smaller projects might just use CI to make sure their code compiles as a minimal check before reviewing a pull request, other projects might have various stages of extensive automated testing.

@@

Semi-structured for build logs means for us that the log text is:
\begin{itemize}
	\item \textbf{implicitly stuructured}, we do not have access to explicit structuring elements or an explicit structure description
	\item \textbf{irregular}, the structure changes without notice \todo{we saw this in our data collection, random newlines}
\end{itemize}

@@

\section{Extractable Information in Build Logs}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
  \caption{Information extractable from build logs}
  \label{fig:build-log-information-draft}
\end{figure}
Continuous integration build logs contain a huge amount of information about the various stages in the CI build they correspond to. In this section we present our model to more precisely describe our notion of a information extractable from a CI build log.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{img/mt-graphics-BuildLogInformation.pdf}
	\caption{Build Log Information}
	\label{fig:build-log-information}
\end{figure}
During our inital exploration \& data collection for the \emph{Failing Build Log Data Set}, wanted to get an impression about the information one would want to extract from a build log. Within the Travis CI build logs we found various different informations possibly interesting to be extracted
\begin{itemize}
	\item \textbf{Travis Build Step} A Travis CI build consists of several build steps defined within the Travis CI configuration language. Within the build log each of these steps are framed by \lstinline{travis_fold:start:<build step name>} and \lstinline{travis_fold:end:<build step name>}.
	      A build step contains
	      \begin{itemize}
		      \item \textbf{Build Step Name} The string Travis uses to identify the step within start \& end statements.
		      \item \textbf{Build Step Output} The output generated during the build step. The textual representation of the \texttt{Build Step Output} is the substring between the start \& end statements.
	      \end{itemize}
	\item \textbf{Travis Timing} Travis can measure the time of specified sections of the build process. It represents those by \lstinline{travis_time:start:<timing section id>} and \lstinline{travis_time:end:<timing section id>:start=<start time>,finish=<finish time>,duration=<duration>}
\end{itemize}

@@

\section{Characteristics of Information Extraction Techniques}
\begin{figure}[h]
	\centering
	\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for an information extraction technique}
	\label{fig:model-ie-technique}
\end{figure}
For this thesis we want to evaluate different Information Extraction Techniques in their application to CI build logs. We focus on techniques that do not attempt or require to parse the structure of a whole build log, but rather focus on extracting just one specific information
A build log information extraction technique (BLIE technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
	\item \textbf{Target}{Extractable Information} each extraction technique \todo{instantiation!} targets a specific information
	\item \textbf{Setup Overhead}{Time}
	\item \textbf{Performance}{Performance} split into learning performance
	\item \textbf{Configuration Scope} for one repository/project, a language/logkind, global
\end{itemize}

@@

\section{Information Extraction Tasks}
\begin{figure}[h]
	\centering
	\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for information extraction task}
	\label{fig:modelt-ie-task}
\end{figure}
aims to describe an information extraction task of a developer or a researcher.
\begin{itemize}
	\item \todo{describe all classes}
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

\section{PROSE Program Synthesis}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Synthesizing regular expression programs by example using PROSE}
  \label{fig:prose-explanation}
\end{figure}
\subsection{Overview}
\emph{Programming by Example (PBE)} is a technique where programs are synthesized according to in and output examples provided by the user. It enables users to create programs without requiring programming knowledge~\cite{mayer2015user}. Applied A to the task of text extraction through regular expressions it relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
We chose to investigate the applicability of generating regular expression programs by example to extracting information from build logs. In this work we refer to our interpretation of this technique as \emph{PBE}. \todo{distinguish PBE for buildlogs}

Generally, we are interested in whether a Programming by Example could replace manually writing regular expressions.
Writing and especially maintaining regular expressions is known to be a difficult, tedious and error-prone task~\cite{michael2019regexes}.
There are two ways to approach this question: from the usability side or the technical side.

When studying the applicability of PBE from the \emph{usability} perspective you run into questions like
\begin{itemize}
	\item Are users faster when using PBE compared to manual regex construction?
	\item Are users more comfortable using PBE over manual regex construction?
	\item Are users confident in the resulting regular expression program from PBE?
	\item Are the resulting extractions more or less accurate when using PBE over manual regex construction?
\end{itemize}
Some of these were already evaluated in existing works about PBE and user interaction~\cite{mayer2015user} \mention{other works?}.
Adequately answering those questions is however highly dependent on the user interface and the corresponding interaction model used during a study.
Developers who are accustomed to crunching on regular expressions in their favorite IDE or Editor might have an adversely negative view on syntesizing programs from examples in an unknow interface.
Mayer et al.~\cite{mayer2015user} presented two system-user interaction models to improve the confidence of users into the programs synthesized from their examples.
We feel such questions of usability can be answered separately from the application domain of build logs.

For our work we chose to look at the applicability of PBE from the \emph{technical} perspective.
We investigate whether current program synthesis techniques are able to learn programs for information extraction tasks in the domain of CI build logs.
Furthermore we gather insights in how PBE has to be applied to yield useable results and perform best.
Our implementation of PBE \todo{which we name technique name x}, is mainly based on the Flash Extract DSL~\cite{flahsextractpaper} of the PROSE library by Microsoft\mention{cite prose website}.
Due to that our technique of text extraction through regular expression programs synthesized is highly influenced by the capabilities of Flash Extract.
In the following we will explain the concept of how we apply PBE to information extraction from CI build logs, the concrete implementation is described in \ref{sec:impl-pbe}.

\subsection*{Configuration / Input / Output}
In/Output examples (I/O examples) are the main driver of Programming by Example.
When extracting information from build logs the \emph{input} is always a whole build log, i.e. the whole text of the build log file without any preprocessing.
In the implementation we did have to apply a few text replacements as described in \ref{sec:impl-preprocessing}.
The \emph{output} is always a substring of the log file text, representing the substring that should be extracted by the syntesized program when given the corresponding input file.
One or multiple I/O examples are needed as configuration for PBEBL, they are used to define substring of a log should be extracted.
For execution PBEBL takes a build file as input and returns a substring of the build file content.

\subsection*{Synthesizing Programs with Flash Extract}
\todo{maybe scrap this section / incorporate last sentence somewhere else, expalnation will mostly be in rw. How to usefully sturcture the stuff above this?}
In the following we want to give an impression on how the PROSE library and specifically the Flash Extract DSL synthesize programs for extracting substrings from text via regular expressions.
\todo{how much can/should we explain here? possibly big parts are already explained in related work: vsa, divied and coquer, witness functions, enumeration and ranking (prose theory paper), flash extract specifics like what functions they have (absolute pos, pre/post regex), rankings specific to ci biulg logs?}
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm. Both are described in \todo{ref rw sections}.

\subsection*{Model instantiation}

\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?, describe regex programs}

\section{Text Similarity}
\begin{figure}[h]
  \centering
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using text similarity}
  \label{fig:text-similarity-explanation}
\end{figure}
\paragraph{Overview}
Information retrieval techniques are more and more prevalently used to extract textual information from unstructured textual sources~\mention{stuff annibale is referencing?}. We describe CI build logs as semi-structured, 
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

### 2019-10-16 thesis/txt/4-tool.tex
[wip] toc corrections
writing about prose technique and some literature notes

\chapter{Our Information Extraction Tool}
\label{sec:implementation}
We implemented prototypes for our three focus information extraction techniques to evaluate them on our collected build logs.
This section will present our unified interface and describe the implementation of each of the techniques.

@@

\section{PROSE Program Synthesis}
The first technique we implemented was the PROSE regular expression program synthesis. As it is based on the PROSE library from Microsoft provided in C\# we also chose C\# for our tool. 
implementation details :)

>>>>

\chapter{Information Extraction Tool}
\label{sec:implementation}
We implemented prototypes for our three focus information extraction techniques to evaluate them on our collected build logs.
This section will present our unified interface and describe the implementation of each of the techniques.

@@

\section{Overall Implementation Remarks}
There are multiple implementation specialities that span over all languages, frameworks and implemented techniques.
The following section discusses preprocessing and escaping of the parsed build log files / example sets, as well as ... \todo{more?}

\subsection{Preprocessing Log Text}
\label{sec:impl-preprocessing}

\section{PROSE Program Synthesis}
\label{sec:impl-pbe}
The first technique we implemented was the PROSE regular expression program synthesis. As it is based on the PROSE library from Microsoft provided in C\# we also chose C\# for our tool. 
implementation details :)
